\sc Cells at War: The Playfulness of Game-Based Learning

### 2023-04-10 thesis_template/paper0904.tex
Updates to discussion

This project has developed an educational video game focused on a single gene disease, which presents students with important cellular and molecular biology concepts in an engaging and interactive manner. Through this study, we have found a way to supplement lecture material with a video game, and examined the impact of game-based learning on student motivation and engagement. This study also shows the immense benefits both students and faculty gain when engaging in a collaborative and interdisciplinary work-integrated learning experience.

@@

@@

>>>>

This project has developed an educational video game focused on a single gene disease, which presents students with important cellular and molecular biology concepts in an engaging and interactive manner. Through this study, we have found a way to enhance the delivery of lecture material with an interactive video game, and examined the impact of game-based learning on student motivation and engagement. This study also highlights the immense benefits both students and faculty gain when engaging in a collaborative and interdisciplinary work-integrated learning experience.

@@

Video games have been around for over 30 years, however, the rate at which technology is evolving has allowed video games to change from short, frenetic, and repetitive forms of engagement, to long, descriptive narratives and story lines \citep{annetta2008video}. Although the results of this study lack quantitative data to support the use of video games in University classrooms, the qualitative feedback provided by students emphasizes the massive potential game-based learning interventions could have at the post-secondary level. \citet{o2009garage} report that static images used in lecture slides are not seemingly effective in conveying dynamic aspects of biological processes. The results from our study provide evidence for this as 69\% of students reported being either somewhat or very prepared if they were to be tested on material using the video game to study, as opposed to studying off of traditional lecture slides. This highlights the need to shift from old fashioned, didactic educational practices, to newer methods that focus on keeping students engaged with the material being presented.

It has also been found that creating a more learner-centered environment in a large introductory biology course by replacing certain material with active learning exercises, can potentially increase engagement, promote critical thinking, and improve student attitudes \citep{moravec2010learn, o2009garage}. Students were asked to report how game based learning interventions could help with following four things; combining ideas from different courses when completing assignments, connecting learning to societal/health related issues, connecting ideas from courses to prior experiences and knowledge, and learning something that changes understanding of a concept. The survey results show an overall increase in the number of students who responded "very much" in all four of these categories, compared to when they were asked how much they were currently doing these things without game-based learning. Video-games have the potential to be powerful educational tools that allow students to connect what they learn through engaging game dynamics, to other aspects of their lives.

It is also interesting to note that the survey results indicated that 41\% of students felt that their coursework emphasized memorizing course material "very much". When asked how a game-based learning intervention could improve motivation to memorize course material, the results show a 14\% decrease in the number of students who responded "very much". This highlights the ability video games have to promote understanding through play, rather than memorization of static textbook and lecture content.

@@

Aside from the benefits that game-based learning has for students utilizing the tool, the educators and students involved in the game design also reap massive benefits. There has been a rising interest in the realm of higher education about "students as partners" and co-created learning and teaching \citep{bibid}
